---
title: Technology Readiness and the Effects of Using RStudio in Teaching a Second Statistics Course at Nueva Vizcaya State University
author: "Orville D. Hombrebueno"
date: "December 16, 2018"
output:
  slidy_presentation: default
  beamer_presentation: default
  ioslides_presentation: default
---
# Rationale
- Technological advancement has led to the emergence of big data (Gould & Çetinkaya-Rundel, 2014) and data science as a field. As a result, the practice of statistics has dramatically changed and has distanced away from statistics education (Finzer, 2013; Wood, Mocko, Everson, Horton, & Velleman, 2018; Zeiffler, Garfield, & Fry, 2018).
- Gould (2010), Horton et al. (2015), Horton (2015) and Hardin et al. (2015) pointed out the importance of developing among students data management skills in introductory and second courses in statistics.
- Franklin et al. (2007) put forth the Guidelines for Assessment and Instruction in Statistics Education (GAISE) Report recommending a framework for statistics education both in the k to 12 and college level. In 2016, the GAISE College Report ASA Revision Committee (2016) revisited the effectiveness of the framework and still found it effective. The framework is now the standard in statistics education in the United States and in many countries that adopted it (Zeiffler et al., 2018). In the GAISE report, one noteworthy recommendation is the “use of technology to explore concepts and analyze data.”

# Statement of the Problem

1. What are the profile variables of the respondents?
2. What are the technology readiness index scores, calculator test scores, and the RStudio test scores of the respondents?
3. What are the technology readiness index scores, the calculator test scores, and the RStudio test scores of the respondents when grouped according to the profile variables?
4. Is there significant difference in the technology readiness index scores, the calculator test scores, and the RStudio test scores of the respondents when grouped according to the profile variables?
5. Is there significant relationship between age and the technology readiness index scores, the calculator test scores, and the RStudio test scores of the respondents?
6. Is there significant relationship between the technology readiness index scores and the RStudio test scores of the respondents?
7. Is there significant difference between the calculator test scores and RStudio test scores of the respondents?

```{r echo=F, message=F}
load("~/R/PhD_1stSem_1819/raw_data.RData")
attach(data)
library(knitr)
library(pander)
```

## Profile Variables of the Respondents

### Gender

```{r echo=F}
a<-as.data.frame(table(Gender),stringsAsFactors = F)
b<-prop.table(table(Gender))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 1:** Frequency Distribution of Gender")
pander(a)
```

### Have Previous Programming Experience

```{r echo=F}
a<-as.data.frame(table(Experienced),stringsAsFactors = F)
b<-prop.table(table(Experienced))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 2:** Frequency Distribution of Have Previous Programming Experience")
pander(a)
```

### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
a<-as.data.frame(table(Preference),stringsAsFactors = F)
b<-prop.table(table(Preference))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 3:** Frequency Distribution of Statistical Computing Tool Preference")
pander(a)
```

### Age

```{r echo=F}
a<-as.data.frame(table(Age),stringsAsFactors = F)
b<-prop.table(table(Age))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
Mean<-mean(Age)
Median<-median(Age)
SD<-sd(Age)
c<-data.frame(Mean,Median,SD)
set.caption("**Table 4:** Frequency Distribution and Summary of Age")
pander(a)
pander(round(c, digits = 2))
```

## Technology Readiness Index Scores, Calculator Test Scores and RStudio Test Scores of the Respondents

### Technology Readiness Index Scores

```{r echo=F}
Min<-min(TRIScore)
Mean<-round(mean(TRIScore),digits = 2)
Max<-max(TRIScore)
SD<-round(sd(TRIScore),digits = 2)
a<-data.frame(Min,Mean,Max,SD)
set.caption("**Table 5:** Summary of Technology Readiness Index Scores")
pander(a)
```

### Calculator Test Scores

```{r echo=F}
Scores<-cut(data$CalcTest, breaks = c(0,49,100), labels = c("Failed","Passed"))
a<-as.data.frame(table(Scores),stringsAsFactors = F)
b<-prop.table(table(Scores))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 6:** Frequency Distribution and Summary of Calculator Test Scores")
Min<-min(CalcTest)
Mean<-mean(CalcTest)
Max<-max(CalcTest)
SD<-sd(CalcTest)
c<-data.frame(Min,Mean,Max,SD)
pander(a)
pander(c)
```

### RStudio Test Scores

```{r echo=F}
Scores<-cut(data$RStudioTest, breaks = c(0,49,100), labels = c("Failed","Passed"))
a<-as.data.frame(table(Scores),stringsAsFactors = F)
b<-prop.table(table(Scores))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 7:** Frequency Distribution and Summary of RStudio Test Scores")
Min<-min(RStudioTest)
Mean<-mean(RStudioTest)
Max<-max(RStudioTest)
SD<-sd(RStudioTest)
c<-data.frame(Min,Mean,Max,SD)
pander(a)
pander(c)
```

## The Technology Readiness Index Scores, Calculator Test Scores, and RStudio Test Scores of the Respondents When Grouped According to the Profile Variables

### Technology Readiness Index Scores

#### Gender

```{r echo=F}
Male<-TRIScore[data$Gender=="Male"]
Female<-TRIScore[data$Gender=="Female"]
Gender<-c("Female","Male")
Min<-c(min(Female),min(Male))
Mean<-round(c(mean(Female),mean(Male)),digits = 2)
Max<-c(max(Female),max(Male))
SD<-round(c(sd(Female),sd(Male)),digits=2)
a<-data.frame(Gender,Min,Mean,Max,SD)
set.caption("**Table 8:** Summary of Technology Readiness Index Scores for Gender")
pander(a)
```

#### Having Previous Programming Experience

```{r echo=F}
No<-TRIScore[data$Prog=="No"]
Yes<-TRIScore[data$Prog=="Yes"]
Experienced<-c("No","Yes")
Min<-c(min(No),min(Yes))
Mean<-round(c(mean(No),mean(Yes)),digits = 2)
Max<-c(max(No),max(Yes))
SD<-round(c(sd(No),sd(Yes)),digits=2)
a<-data.frame(Experienced,Min,Mean,Max,SD)
set.caption("**Table 9:** Summary of Technology Readiness Index Scores for Having Previous Programming Experience")
pander(a)
```

#### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
Calculator<-TRIScore[data$RStudio=="Calculator"]
RStudio<-TRIScore[data$RStudio=="RStudio"]
Preference<-c("Calculator","RStudio")
Min<-c(min(Calculator),min(RStudio))
Mean<-round(c(mean(Calculator),mean(RStudio)),digits = 2)
Max<-c(max(Calculator),max(RStudio))
SD<-round(c(sd(Calculator),sd(RStudio)),digits=2)
a<-data.frame(Preference,Min,Mean,Max,SD)
set.caption("**Table 10:** Summary of Technology Readiness Index for Statistical Computing Tool Preference")
pander(a)
```

### Calculator Test Scores

#### Gender

```{r echo=F}
Male<-CalcTest[data$Gender=="Male"]
Female<-CalcTest[data$Gender=="Female"]
Gender<-c("Female","Male")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Gender,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Female),min(Male))
Mean<-round(c(mean(Female),mean(Male)),digits = 2)
Max<-c(max(Female),max(Male))
SD<-round(c(sd(Female),sd(Male)),digits=2)
a<-data.frame(Gender,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 11:** Frequency Distribution and Summary of Calculator Test Scores for Gender")
pander(a)
```

#### Having Previous Programming Experience

```{r echo=F}
No<-CalcTest[data$Prog=="No"]
Yes<-CalcTest[data$Prog=="Yes"]
Experienced<-c("No","Yes")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Prog,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(No),min(Yes))
Mean<-round(c(mean(No),mean(Yes)),digits = 2)
Max<-c(max(No),max(Yes))
SD<-round(c(sd(No),sd(Yes)),digits=2)
a<-data.frame(Experienced,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 12:** Frequency Distribution and Summary of Calculator Test Scores for Having Previous Programming Experience")
pander(a)
```

#### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
Calculator<-CalcTest[data$RStudio=="Calculator"]
RStudio<-CalcTest[data$RStudio=="RStudio"]
Preference<-c("Calculator","RStudio")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$RStudio,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Calculator),min(RStudio))
Mean<-round(c(mean(Calculator),mean(RStudio)),digits = 2)
Max<-c(max(Calculator),max(RStudio))
SD<-round(c(sd(Calculator),sd(RStudio)),digits=2)
a<-data.frame(Preference,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 13:** Frequency Distribution and Summary of Calculator Test Scores for Statistical Computing Tool Preference")
pander(a)
```

### RStudio Test Scores

#### Gender

```{r echo=F}
Male<-RStudioTest[data$Gender=="Male"]
Female<-RStudioTest[data$Gender=="Female"]
Gender<-c("Female","Male")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Gender,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Female),min(Male))
Mean<-round(c(mean(Female),mean(Male)),digits = 2)
Max<-c(max(Female),max(Male))
SD<-round(c(sd(Female),sd(Male)),digits=2)
a<-data.frame(Gender,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 14:** Frequency Distribution and Summary of RStudio Test Scores for Gender")
pander(a)
```

#### Having Previous Programming Experience

```{r echo=F}
No<-RStudioTest[data$Prog=="No"]
Yes<-RStudioTest[data$Prog=="Yes"]
Experienced<-c("No","Yes")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Prog,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(No),min(Yes))
Mean<-round(c(mean(No),mean(Yes)),digits = 2)
Max<-c(max(No),max(Yes))
SD<-round(c(sd(No),sd(Yes)),digits=2)
a<-data.frame(Experienced,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 15:** Frequency Distribution and Summary of RStudio Test Scores for Having Previous Programming Experience")
pander(a)
```

#### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
Calculator<-RStudioTest[data$RStudio=="Calculator"]
RStudio<-RStudioTest[data$RStudio=="RStudio"]
Preference<-c("Calculator","RStudio")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$RStudio,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Calculator),min(RStudio))
Mean<-round(c(mean(Calculator),mean(RStudio)),digits = 2)
Max<-c(max(Calculator),max(RStudio))
SD<-round(c(sd(Calculator),sd(RStudio)),digits=2)
a<-data.frame(Preference,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 16:** Frequency Distribution and Summary of RStudio Test Scores for Statistical Computing Tool Preference")
pander(a)
```

## Significant Difference in the Technology Readiness Index Scores, the Calculator Test Scores, and the RStudio Test Scores of the Respondents when Grouped According to the Profile Variables

### Technology Readiness Index Scores

There are no significant difference.

### Calculator Test Scores

There are no significant difference.

### RStudio Test Scores

The Mann-Whitney test

```{r echo=F}
Female<-RStudioTest[data$Gender=="Female"]
Male<-RStudioTest[data$Gender=="Male"]
pander(wilcox.test(Male,Female,alternative = "g"))
```

## Significant Relationship Between Age and the Technology Readiness Index Scores, the Calculator Test Scores and the RStudio Test Scores of the Respondents

### Technology Readiness Index Scores and Age

No significant relationship.

### Calculator Test Scores and Age

No Significant relationship.

### RStudio Test Scores and Age

No significant relationship.

## Significant Relationship Between the Technology Readiness Index Scores and the RStudio Test Scores of the Respondents

There is no significant relationship.

## Significant Difference Between the Calculator Test Scores and Rstudio Test Scores Of The Respondents

There is no significant difference.