---
output:
  word_document:
    reference_docx: Research_CIA_my_style.docx
bibliography: biblio.bib
csl: apa.csl
---

# Introduction

## Rationale

The practice of statistics is computational. Statistical computing software are mainly used to do statistical computing. With the emergence of data science as a field due to big data, machine learning and powerful computers that are no longer expensive all brought about by the 4th industrial revolution, it is very fitting to consider how statistics education should adapt to these changes and be relevant in the practice of statistics. In this regard, in 2005, @franklin07 put forth the Guidelines for Assessment and Instruction in Statistics Education (GAISE) Report recommending a framework for statistics education both in the k to 12 and college level. In 2016, the @GCR16 revisited the effectiveness of the framework and still found it effective. The framework is now the standard in statistics education in the United States and in many countries that adopted it [@zieffler18].

In the GAISE report, one noteworthy recommendation is "the use of technology to explore concepts and analyze data." Studies have shown that the use of technology can really improve statistics education; equipping learners with relevant data skills and effective powerful tools in this era where data is very much abundant [@doi16; @chance06; @chance07; @harraway12; @CR17; @stander17].  However, when it comes to software used in doing statistical computing and teaching statistics, there is no single statistical computing tool that fits all statistical tasks [@mcnamara18]. Nonetheless, introductory statistics students should be taught a common statistical computing software such as SAS, SPSS, or R [@R], enthusing them to continuously learn statistics technology since statistical tools are diverse and eventually evolve through time [@gould18a]. Moreover, @gould10, @horton15a, @horton15 and @hardin15 pointed out the importance of developing among students -- with the use of statistical computing software -- data management skills in introductory and second courses in statistics.

In our country, leading universities are addressing this issue and have already integrated the use of technology in their curriculum for statistics education. The University of the Philippines for example uses a number of software in its introductory statistics courses and statistics courses (eg. R with RStudio, Python, SAS, SPSS, Stata, MS Excel, QGIS, ArcGIS, Gephi, yEd Graph Editor, and more). Most of the software used are opensource. 

In the province, some universities have acquired SPSS to teach statistics courses. On the other hand, some still uses calculators to do and teach statistical computing. Unfortunately, both setup do not lessen the gap between statistics education and statistical practice. In reality, most institutions to which graduates from these universities get employed cannot afford SPSS. One can choose MS Excel as an alternative, however its functions are limited [@biehler13]. 

These developments gave way to the emergence of data science as a field.  As a result, the practice of statistics has dramatically changed and has distanced away from statistics education [@finzer13; @wood18; @zieffler18].  Nonetheless, some measures are already in place to lessen the gap between statistical practice and statistics education.   

Moreover, @gould10, @horton15a, @horton15 and @hardin15 pointed out the importance of developing among students data management skills in introductory and second courses in statistics.


Reproduciblequite

The gap between statistical practice and statistics education.

barriers in using technology (price)
 

R and RStudio

Efforts in the Philippines

## Statement of the Problem

1. What are the profile variables of the respondents?

2. What are the technology readiness index scores, the calculator test scores, and the RStudio test scores of the respondents when grouped according to the profile variables?

3. Is there significant difference in the calculator test scores and the RStudio test scores of the respondents when grouped according to the profile variables?

4. Is there significant relationship between age and the calculator test scores, and the RStudio test scores of the respondents?

5. Is there significant difference between the calculator test scores and RStudio test scores of the respondents?

## Statement of the Hypothesis

1. Is there significant difference between the calculator test scores and RStudio test scores of the respondents?

2. the 

## Significance of the Study



## Research Framework



## Scope 


## Definition of Terms



# Review of Related Literature and Studies

The choice of which software to use in teaching statistics can be quite a challenge given a lot of things to consider. It can be a problem similar to bridging the gap between the practice of statistics and statistics education. Tools for learning statistics can be used easily by starters opposite to when using tools for doing statistics [@gould18a]. However,

# Methodology



## Research Design



## Research Environment



## Respondents



## Research Instruments



## Data Gathering Procedure



## Data Analysis



# Results and Discussion

```{r echo=F, message=F}
load("~/R/PhD_1stSem_1819/raw_data.RData")
attach(data)
library(knitr)
library(pander)
```

## Profile Variables of the Respondents

The profile variables of the respondents considered in the study are gender: male or female, have previous programming experience: yes or no, statistical computing preference: calculator or RStudio, and age.

### Gender

Most of those who enroll BSEd mathematics are female. It is no surprise that among the respondents, 62.50% are female and 37.50% are male. Table 1 shows the frequency distribution of gender.

```{r echo=F}
a<-as.data.frame(table(Gender),stringsAsFactors = F)
b<-prop.table(table(Gender))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 1:** Frequency Distribution of Gender")
pander(a)
```

### Have Previous Programming Experience

Among the respondents, half have previos programming experience while the other half have none. Table 2 shows the frequency distribution for have previous programming experience. It seems that some students still have no programming experience when they take their second course in statistics.

```{r echo=F}
a<-as.data.frame(table(Experienced),stringsAsFactors = F)
b<-prop.table(table(Experienced))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 2:** Frequency Distribution of Have Previous Programming Experience")
pander(a)
```

### Statistical Computing Tool Preference: Calculator or RStudio

Having learned the basics of using calculator and RStudio for statistical computing and took the examination using calculator and RStudio and then knowing their scores, half of the respondents still preferred to use calculator while the other half now preferred to use RStudio. Table 3 shows the frequency distribution for statistical computing tool preference. Among those who preferred RStudio, one said "it is easy". Among those who preferred calculator, one found using RStudio "complicated" and that using calculator is "comfortable". Table 3 shows the frequency distribution of statistical computing tool preference.

```{r echo=F}
a<-as.data.frame(table(Preference),stringsAsFactors = F)
b<-prop.table(table(Preference))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 3:** Frequency Distribution of Statistical Computing Tool Preference")
pander(a)
```

### Age

Half of the respondents are 21 years old, two are 22 years old, one is 24 years old and one is 39 years old. Table 4 is the frequency distribution of age.

```{r echo=F}
a<-as.data.frame(table(Age),stringsAsFactors = F)
b<-prop.table(table(Age))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 4:** Frequency Distribution of Age")
pander(a)
```

## Calculator Test Scores and RStudio Test Scores of the Respondents

There are two sets of examination scores: the calculator test scores and RStudio test scores.

### Calculator Test Scores

The result of the calculator test shows that 62.50% failed and 37.50% passed. The mean score is 46.25% which is not even passing. The exam result is not remarkable. Table 5 shows the frequency distribution and summary of calculator test scores.

```{r echo=F}
Scores<-cut(data$CalcTest, breaks = c(0,49,100), labels = c("Failed","Passed"))
a<-as.data.frame(table(Scores),stringsAsFactors = F)
b<-prop.table(table(Scores))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 5:** Frequency Distribution and Summary of Calculator Test Scores")
Min<-min(CalcTest)
Mean<-mean(CalcTest)
Max<-max(CalcTest)
SD<-sd(CalcTest)
c<-data.frame(Min,Mean,Max,SD)
pander(a)
pander(c)
```

### RStudio Test Scores

The result of the RStudio test shows that 37.50% passed and 62.50% failed. The mean score is 54.50 which is passing. Compared to the calculator test scores, the RStudio test scores seem to have a better result. Table 6 shows the frequency distribution and summary of RStudio test scores.

```{r echo=F}
Scores<-cut(data$RStudioTest, breaks = c(0,49,100), labels = c("Failed","Passed"))
a<-as.data.frame(table(Scores),stringsAsFactors = F)
b<-prop.table(table(Scores))
a$Percentage<-round(100*c(b),digits=2)
a[nrow(a)+1,]<-list("Total", 8, 100)
set.caption("**Table 6:** Frequency Distribution and Summary of RStudio Test Scores")
Min<-min(RStudioTest)
Mean<-mean(RStudioTest)
Max<-max(RStudioTest)
SD<-sd(RStudioTest)
c<-data.frame(Min,Mean,Max,SD)
pander(a)
pander(c)
```

## Calculator Test Scores and RStudio Test Scores of the Respondents When Grouped According to the Profile Variables

The calculator test scores and the RStudio test scores of the respondents are grouped according to the profile variables gender, having previous programming experience and statistical computing tool preference.

### Calculator Test Scores According to Gender

Among the female respondents, only one passed the test. Among the male respondents, only one failed the test. The mean score for female respondents is 37 which is not passing. The mean score for male respondents is 61.67 which is passing. Table 7 shows the frequency distribution and summary of calculator test scores according to gender. The calculator test scores of male respondents is better than the scores of female respondents.

```{r echo=F}
Male<-CalcTest[data$Gender=="Male"]
Female<-CalcTest[data$Gender=="Female"]
Gender<-c("Female","Male")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Gender,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Female),min(Male))
Mean<-round(c(mean(Female),mean(Male)),digits = 2)
Max<-c(max(Female),max(Male))
SD<-round(c(sd(Female),sd(Male)),digits=2)
a<-data.frame(Gender,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 7:** Frequency Distribution and Summary of Calculator Test Scores According to Gender")
pander(a)
```

### RStudio Test Scores According to Gender

Among the female respondents, the number of respondents who failed the test is one respondent greater than the number of respondents who passed the test. All of the male respondents passed the test. The mean score for female respondents is 43.20 which is not passing. The mean score for male respondents is 73.33 which is passing. Table 8 shows the frequency distribution and summary of RStudio test scores according to gender. The RStudio test scores of male respondents seem to be better than the scores of female respondents.

```{r echo=F}
Male<-RStudioTest[data$Gender=="Male"]
Female<-RStudioTest[data$Gender=="Female"]
Gender<-c("Female","Male")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Gender,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Female),min(Male))
Mean<-round(c(mean(Female),mean(Male)),digits = 2)
Max<-c(max(Female),max(Male))
SD<-round(c(sd(Female),sd(Male)),digits=2)
a<-data.frame(Gender,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 8:** Frequency Distribution and Summary of RStudio Test Scores According to Gender")
pander(a)
```

### Calculator Test Scores According to Having Previous Programming Experience

Half of the respondents who have previous programming experience failed the test and the other half passed. Only one respondent who have previous programming experience passed the test.

```{r echo=F}
No<-CalcTest[data$Prog=="No"]
Yes<-CalcTest[data$Prog=="Yes"]
Experienced<-c("No","Yes")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Prog,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(No),min(Yes))
Mean<-round(c(mean(No),mean(Yes)),digits = 2)
Max<-c(max(No),max(Yes))
SD<-round(c(sd(No),sd(Yes)),digits=2)
a<-data.frame(Experienced,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 9:** Frequency Distribution and Summary of Calculator Test Scores According to Having Previous Programming Experience")
pander(a)
```

### RStudio Test Scores According to Having Previous Programming Experience



```{r echo=F}
No<-RStudioTest[data$Prog=="No"]
Yes<-RStudioTest[data$Prog=="Yes"]
Experienced<-c("No","Yes")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$Prog,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(No),min(Yes))
Mean<-round(c(mean(No),mean(Yes)),digits = 2)
Max<-c(max(No),max(Yes))
SD<-round(c(sd(No),sd(Yes)),digits=2)
a<-data.frame(Experienced,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 10:** Frequency Distribution and Summary of RStudio Test Scores According to Having Previous Programming Experience")
pander(a)
```

#### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
Calculator<-CalcTest[data$RStudio=="Calculator"]
RStudio<-CalcTest[data$RStudio=="RStudio"]
Preference<-c("Calculator","RStudio")
b<-cut(CalcTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$RStudio,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Calculator),min(RStudio))
Mean<-round(c(mean(Calculator),mean(RStudio)),digits = 2)
Max<-c(max(Calculator),max(RStudio))
SD<-round(c(sd(Calculator),sd(RStudio)),digits=2)
a<-data.frame(Preference,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 13:** Frequency Distribution and Summary of Calculator Test Scores for Statistical Computing Tool Preference")
pander(a)
```

#### Statistical Computing Tool Preference: Calculator or RStudio

```{r echo=F}
Calculator<-RStudioTest[data$RStudio=="Calculator"]
RStudio<-RStudioTest[data$RStudio=="RStudio"]
Preference<-c("Calculator","RStudio")
b<-cut(RStudioTest, breaks = c(0,49,100),labels = c("Failed","Passed"))
c<-c(table(data$RStudio,b))
Failed<-c[1:2]
Passed<-c[3:4]
Min<-c(min(Calculator),min(RStudio))
Mean<-round(c(mean(Calculator),mean(RStudio)),digits = 2)
Max<-c(max(Calculator),max(RStudio))
SD<-round(c(sd(Calculator),sd(RStudio)),digits=2)
a<-data.frame(Preference,Failed,Passed,Min,Mean,Max,SD)
set.caption("**Table 16:** Frequency Distribution and Summary of RStudio Test Scores for Statistical Computing Tool Preference")
pander(a)
```

## Significant Difference in the Technology Readiness Index Scores, the Calculator Test Scores, and the RStudio Test Scores of the Respondents when Grouped According to the Profile Variables

### Technology Readiness Index Scores

There are no significant difference.

### Calculator Test Scores

There are no significant difference.

### RStudio Test Scores

The Mann-Whitney test

```{r echo=F}
Female<-RStudioTest[data$Gender=="Female"]
Male<-RStudioTest[data$Gender=="Male"]
pander(wilcox.test(Male,Female,alternative = "g"))
```

## Significant Relationship Between Age and the Technology Readiness Index Scores, the Calculator Test Scores and the RStudio Test Scores of the Respondents

### Technology Readiness Index Scores and Age

No significant relationship.

### Calculator Test Scores and Age

No Significant relationship.

### RStudio Test Scores and Age

No significant relationship.

## Significant Relationship Between the Technology Readiness Index Scores and the RStudio Test Scores of the Respondents

There is no significant relationship.

## Significant Difference Between the Calculator Test Scores and Rstudio Test Scores Of The Respondents

There is no significant difference.

# Reference
